---
title: "PTCF project"
author: "Huanhuan Shi"
date: "5/9/2022"
output: 
    pdf_document
---


```{r,include= FALSE, echo=FALSE,message=FALSE,warning=FALSE}
###load the packages and  data
library(dplyr)
library(psych)
library(lme4)
library(ggplot2)
library(lmerTest)
library(broom)
library(broom.mixed)
library(knitr)
d <- read.csv("PTCF.csv")
```



```{r, include=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
##age center
d$AgeCenter<-d$age-mean(d$age)
head(d)
summary(d)
```

```{r,include=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
#data for priming 
d_Prime<-d%>%
  filter(task=="priming")
d_Story<-d%>%
  filter(task=="storyretelling")
head(d_Story)
d_Spoken<-d%>%
  filter(modality=="spoken")
head(d_Spoken)
d_Written<-d%>%
  filter(modality=="written")
```

```{r,include=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
##get the mean for each condition
###mean for PrimeOral
mean_Prime<-mean(d_Prime$accuracy)
mean_Prime

###mean for PrimeWritten
mean_Story<-mean(d_Story$accuracy)
mean_Story

###mean for StorySpoken
mean_Spoken<-mean(d_Spoken$accuracy)
mean_Spoken


###mean for PrimeOral
mean_Written<-mean(d_Written$accuracy)
mean_Written
```


```{r, include=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
## summarize with group_by task
d_task<-d %>% 
  group_by(task) %>%
  summarize(meanaccuracy=mean(accuracy),
            N = length(accuracy),
            mean = mean(accuracy),
            sd   = sd(accuracy),
            se   = sd / sqrt(N))
            
d_task

```


```{r, include=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
##group by modality
d_modality<-d %>% 
  group_by(modality) %>%
  summarize(meanaccuracy=mean(accuracy),
            N = length(accuracy),
            mean = mean(accuracy),
            sd   = sd(accuracy),
            se   = sd / sqrt(N))

d_modality
```


```{r,include=FALSE, echo=FALSE,message=FALSE,warning=FALSE}
# group by condition
d_condition<-d %>% 
  group_by(condition) %>%
  summarize(meanaccuracy=mean(accuracy),
            N = length(accuracy),
            mean = mean(accuracy),
            sd   = sd(accuracy),
            se   = sd / sqrt(N))

head(d_condition)
summary(d_condition)
```



```{r, include= FALSE,echo=FALSE,message=FALSE,warning=FALSE}
head(d)
d$M<-ifelse(d$modality=="spoken",-0.5,0.5)
d$T<-ifelse(d$task=="priming",-0.5,0.5)
lm<-glmer(accuracy ~ (1|subject) +(1|picture)+ AgeCenter+ T*M, data=d, family = binomial)
summary(lm)
```

```{r coefficiency, include= FALSE, echo=FALSE,message=FALSE,warning=FALSE}
##get the paprameter in the model
coeff <- tidy(lm)
coeff
```
**Project Description**

 This project examined how response modality (spoken vs. written) and elicitaton context (prming vs.storyretelling) affected adolescents' production of past tense counterfactual sentences(PTCF). This is an overview of the data.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
head(d)
```

This is a summary of the data.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
summary(d)
```


The descriptive data was presented in Table 1.

```{r tables for all conditions, echo=FALSE,message=FALSE,warning=FALSE}
kable(d_condition, col.names=c("Conditions","Accuracy","N","mean","sd","se"),align = "cccccc",
      caption="Table 1. Descriptive data for the four conditions")
```

Figure 1. presented the same results as Table 1.

```{r,fig.align = 'center',out.width= '80%',fig.cap='Fiture 1. Mean proportion of the canonical PTCF sentences in each condition.', echo=FALSE,message=FALSE,warning=FALSE}
##with color narrow bar
ggplot(d_condition, aes(x=condition, y=mean, fill=condition)) + 
  geom_bar(stat="identity", width = .8) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1, position=position_dodge(.9))+ xlab("Conditions") + ylim(0,1) + ylab("Proportion of canonical PTCF ")+ guides(fill="none")+theme_bw()+ theme(axis.title=element_text(face="bold",size=12),axis.text.x=element_text(face="bold",size=10),axis.text.y=element_text(size=10),strip.text.x=element_text(face="bold",size=12),legend.text=element_text(face="bold",size=15))


###no lines in the background
ggplot(d_condition, aes(x=condition, y=mean, fill=condition)) + 
  geom_bar(stat="identity", width = 0.8) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.9)) + 
  xlab("Conditions") + 
  ylim(0, 1) + 
  ylab("Proportion of Canonical PTCF Sentences") + 
  guides(fill="none") + 
  theme_bw() + 
  theme(
    axis.title = element_text(face="bold", size=12),
    axis.text.x = element_text(face="bold", size=10),
    axis.text.y = element_text(size=10),
    strip.text.x = element_text(face="bold", size=12),
    legend.text = element_text(face="bold", size=15),
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank()   # Remove minor grid lines
  )

```

**Model output**

We run a binomial mixed-effects regression with task, modality and age as fixed effects and subjects and trials as random effects. There was a main effect of task (*$\beta$* = `r round(coeff[3,4], digits=2)`,*p* < .001).The participants produced significantly more PTCF sentences in the priming tasks (*mean* = `r round(mean_Prime, digits=2)`) than in the story retelling tasks (*mean* = `r round(mean_Story, digits=2)`). See Figure 2.

We also found a main effect of modality (*$\beta$* = `r round(coeff[4,4], digits=2)`,*p* < .001), which revealed that the participants used more PTCF setences in spoken (*mean* = `r round(mean_Spoken, digits=2)`) than in written production (*mean* = `r round(mean_Written, digits=2)`). There was no effect of age (*$\beta$* = `r round(coeff[3,4], digits=2)`,*p* = `r round(coeff[2,7], digits=2)`), and no interaction between task and modality (*$\beta$* = `r round(coeff[5,4], digits=2)`, *p* = `r round(coeff[5,7], digits=2)`). See Figure 3.


```{r,fig.align = 'center',out.width= '80%',fig.cap='Mean proportion of canonical PTCF sentences in each task.',echo=FALSE,message=FALSE,warning=FALSE}
##figure by task
ggplot(d_task, aes(x=task, y=mean, fill=task)) + 
  geom_bar(stat="identity", width = .8) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1, position=position_dodge(.9))+ xlab("Tasks") + ylim(0,1) + ylab("Proportion of Canonical PTCF Sentences")+ guides(fill="none")+theme_bw()+ theme(axis.title=element_text(face="bold",size=12),axis.text.x=element_text(face="bold",size=10),axis.text.y=element_text(size=12),strip.text.x=element_text(face="bold",size=10),legend.text=element_text(face="bold",size=12))


###no background lines
p3 <- ggplot(d_task, aes(x=task, y=mean, fill=task)) + 
  geom_bar(stat="identity", width = 0.6) +  # Made bars even narrower by reducing width
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.9)) + 
  xlab("Tasks") + 
  ylim(0, 1) + 
  ylab("Proportion of Canonical PTCF Sentences") + 
  guides(fill="none") + 
  theme_bw() + 
  theme(
    axis.title = element_text(face="bold", size=12),
    axis.title.y = element_text(size=12),  # Make y-axis label font smaller
    axis.text.x = element_text(face="bold", size=10),
    axis.text.y = element_text(size=12),
    strip.text.x = element_text(face="bold", size=12),
    legend.text = element_text(face="bold", size=12),
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank()   # Remove minor grid lines
  )

# Save the plot with increased height
ggsave("Figure3.png", plot = p3, width = 4, height = 4,dpi = 500)
```


```{r,fig.align = 'center',out.width= '80%',fig.cap= 'Mean proportion of canonical PTCF sentences in each modality.', echo=FALSE,message=FALSE,warning=FALSE}
##figure by modality
ggplot(d_modality, aes(x=modality, y=mean, fill=modality)) + 
  geom_bar(stat="identity", width = .8) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1, position=position_dodge(.9))+ xlab("Modalities") + ylim(0,1) + ylab("Proportion of canonical PTCF ")+ guides(fill="none")+theme_bw()+ theme(axis.title=element_text(face="bold",size=12),axis.text.x=element_text(size=10),axis.text.y=element_text(size=12),strip.text.x=element_text(face="bold",size=10),legend.text=element_text(face="bold",size=12))


###no lines in the backgound
p4 <- ggplot(d_modality, aes(x=modality, y=mean, fill=modality)) + 
  geom_bar(stat="identity", width = 0.6) +  # Made bars narrower by reducing width
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.9)) + 
  xlab("Modalities") + 
  ylim(0, 1) + 
  ylab("Proportion of Canonical PTCF Sentences") + 
  guides(fill="none") + 
  theme_bw() + 
  theme(
    axis.title = element_text(face="bold", size=12),
    axis.text.x = element_text(face="bold", size=10),
    axis.text.y = element_text(size=12),
    strip.text.x = element_text(face="bold", size=10),
    legend.text = element_text(face="bold", size=12),
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank()   # Remove minor grid lines
  )

# Save the plot with narrower dimensions
ggsave("Figure4.png", plot = p4, width = 4, height = 4,dpi = 500)

```


